Automated iterative forecasting for the Portal Project

### 2018-02-05 manuscript/portal_iterative_forecasting_ms.Rmd
Update archiving section to include external archiving
Closes #172

Co-authored-by: Morgan Ernest <morgan@weecology.org>

Publicly archiving forecasts before new data is collected allows the field to assess, compare, and build on forecasts made by different groups [@mcgill2012; @tredennick2016; @dietze2017; @harris2018] (Figure 1). Archiving serves as a form of pre-registration for model predictions, helping facilitate unbiased interpretation of model performance. We wanted our archives to not only be accessible to others but to also be a permanent record whose continued existence was not dependent on any particular researcher or funding stream. Storing forecast outputs in our forecasting GitHub repository was not sufficient for archival purposes because repositories can be deleted.  We explored three major repositories for archiving our forecasts: FigShare (https://figshare.com/), Zenodo (https://zenodo.org/), and Open Science Framework (https://osf.io/). While all three repositories allowed for easy manual submissions (i.e., a human uploading files after each forecast), automating this process was substantially more difficult. Various combinations of repositories, APIs (i.e., interfaces that allow automated ways interacting with these websites) and associated R packages had issues with: 1) integrating authorization with continuous integration; 2) automatically making archived files public; 3) adding new files to an existing location; or 4) automatically permanently archiving the files. Our eventual solution was to leverage the GitHub-Zenodo integration (https://guides.github.com/activities/citable-code/) and automatically push forecasts to the GitHub repository from the CI server. The GitHub-Zenodo integration is designed to automatically create versioned archives of GitHub repositories. There is an existing one-time process for linking our forecasting repository on  GitHub with Zenodo. Once this link is created, each time a new forecast is created, our pipeline adds the new forecasts to the GitHub repository and uses the GitHub API to create a new "release" of our repository. This triggers the Zenodo-GitHub integration, which automatically archives the resulting forecasts and the code that generated them under a top-level DOI that refers to all archived forecasts ([https://doi.org/10.5281/zenodo.833438](https://doi.org/10.5281/zenodo.833438)). Through this process, we automatically archive every forecast made with a documented history of the archive. While this approach is functional because everything in the repository is archived when a new forecast is made, these archives have complicated folder structures, making it more harder for users to find and access the forecasting results. Finding simpler approaches to automated archiving would increase the transparency of the forecase results and their ease of use by external researchers.

>>>>

Publicly archiving forecasts before new data is collected allows the field to assess, compare, and build on forecasts made by different groups [@mcgill2012; @tredennick2016; @dietze2018; @harris2018] (Figure 1). Archiving serves as a form of pre-registration for model predictions, because the forecasts cannot be modified once the data to assess them has been collected. This helps facilitate an unbiased interpretation of model performance. To serve this role archives should be publicly accessible and be a permanent record that cannot be changed or deleted. The second criteria means that GitHub is not sufficient for archival purposes because repositories can changed or deleted [@bergman2012].  We explored three major repositories for archiving forecasts: FigShare (https://figshare.com/), Zenodo (https://zenodo.org/), and Open Science Framework (https://osf.io/). While all three repositories allowed for easy manual submissions (i.e., a human uploading files after each forecast), automating this process was substantially more difficult. Various combinations of repositories, APIs (i.e., interfaces for automatically interacting with these websites) and associated R packages had issues with: 1) integrating authorization with continuous integration; 2) automatically making archived files public; 3) adding new files to an existing location; or 4) automatically permanently archiving the files. Our eventual solution was to leverage the GitHub-Zenodo integration (https://guides.github.com/activities/citable-code/) and automatically push forecasts to a GitHub repository from the CI server and release them via the GitHub API. The GitHub-Zenodo integration is designed to automatically create versioned archives of GitHub repositories. We created a repository for storing forecasts (https://github.com/weecology/forecasts) and linked this repository with Zenodo (a one-time manual process). Each time a new forecast is created, our pipeline adds the new forecasts to the GitHub repository and uses the GitHub API to create a new "release" for that repository. This triggers the Zenodo-GitHub integration, which automatically archives the resulting forecasts under a top-level DOI that refers to all archived forecasts ([https://doi.org/10.5281/zenodo.839580](https://doi.org/10.5281/zenodo.839580)). Through this process, we automatically archive every forecast made with a documented time-stamp for every forecast in the archive. In addition, we also archive the full state of the repository for modeling and forecasting ([https://doi.org/10.5281/zenodo.833438](https://doi.org/10.5281/zenodo.833438)). This ensures that every forecast is fully reproducible since the exact code used to generate every forecast is preserved. Early forecasts from this system are archived as part of this modeling and forecasting code archive.
